---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(str = strOptions(strict.width = "cut"))
```

# gleifr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/m-muecke/gleifr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/gleifr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

gleifr is a minimal R client for the [GLEIF](https://www.gleif.org) API.
A major challenge when dealing with financial data is mapping entities across different data sources.
The [Legal Entity Identifier (LEI)](https://www.gleif.org/en/organizational-identity/lei-vlei/the-legal-entity-identifier-lei) is a 20-character, alphanumeric code that uniquely identifies legally distinct entities that engage in financial transactions.
The Global Legal Entity Identifier Foundation (GLEIF) maintains the global LEI system and exposes it through a free, public API.

gleifr wraps that API so you can, from R:

- look up the reference data behind any LEI, or search for LEIs by name, jurisdiction, or full text,
- traverse corporate hierarchies via direct and ultimate parent and child relationships,
- map LEIs to other identifiers such as ISIN, BIC, and MIC,
- and access supporting reference data like regions and accredited issuers (Local Operating Units).

## Installation

You can install the development version of gleifr from [GitHub](https://github.com/m-muecke/gleifr) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/gleifr")
```

## Usage

```{r}
library(gleifr)
```

### Looking up a record by LEI

Pass a LEI to `lei_records()` to retrieve its full reference data.
By default the result is returned as a tidy long-format data frame with one attribute per row:

```{r}
records <- lei_records("001GPB6A9XPE8XJICC14")
str(records)
```

### Searching for records

When called without an `id`, `lei_records()` searches the LEI database.
You can filter by legal name, jurisdiction, status, or run a full-text query:

```{r}
#| eval: false
# search by legal name
lei_records(legal_name = "Deutsche Bank")

# combine filters
lei_records(jurisdiction = "DE", status = "ACTIVE")
```

By default all matching pages are fetched automatically.
Use `page_size` and `page_number` to page through results manually, and `...` to pass any additional GLEIF filter, e.g. `"filter[entity.legalAddress.country]" = "DE"`.

### Corporate relationships

Traverse ownership hierarchies with `lei_parents()` and `lei_children()`:

```{r}
#| eval: false
# direct parent (use type = "ultimate" for the ultimate parent)
lei_parents("529900W18LQJJN6SJ336")

# direct children
lei_children("529900W18LQJJN6SJ336")
```

### Identifier mappings

`lei_mapping()` downloads the latest bulk relationship file that maps every LEI to another identifier type: ISIN, BIC, MIC, or OpenCorporates ID (`"oc"`):

```{r}
mapping <- lei_mapping("isin")
head(mapping)
```

To fetch the ISINs for a single entity instead of the full file, use `lei_isins()`:

```{r}
#| eval: false
lei_isins("529900W18LQJJN6SJ336")
```

## Caching

API responses can be cached locally to avoid repeated downloads.
Caching is opt-in and controlled through options:

```{r}
#| eval: false
# enable caching (responses are kept for 1 day by default)
options(gleifr.cache = TRUE)

# customize the maximum age in seconds
options(gleifr.cache_max_age = 3600)

# inspect or clear the cache
lei_cache_dir()
lei_cache_clear()
```
