---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(str = strOptions(strict.width = "cut"))
```

# gleifr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/m-muecke/gleifr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/gleifr/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/gleifr)](https://CRAN.R-project.org/package=gleifr)
<!-- badges: end -->

gleifr is a minimal R client for the [GLEIF](https://www.gleif.org) API.
A major challenge when dealing with financial data is the mapping of entities across different data sources.
Especially when dealing with legal entities, the Legal Entity Identifier (LEI) can be used to uniquely identify entities.

## Installation

You can install the development version of gleifr from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/gleifr")
```

## Usage

Currently the download of the lei mapping data and the retrieval of records by LEI is supported:

```{r}
library(gleifr)

# fetch the latest LEI mapping data (single file with all mappings)
mapping <- lei_mapping("isin")
head(mapping)

# fetch LEI records by a given LEI identifier
records <- lei_records("001GPB6A9XPE8XJICC14")
str(records)
```

## Related work

- [gleifr](https://github.com/Financial-Times/gleifr) - R package to support analysis of GLEIF data
